---
title: "KThomas_HW5"
author: "Kyle Thomas"
date: "September 27, 2017"
output: html_document
---

# Homework 5

## Question 1


```{r}
#set the path
setwd("C:\\Users\\kthomas\\Documents\\SMU\\msds6306 - doing data science\\unit 5\\homework\\Download")

#read in the data to a data frame
dfNames= read.table("yob2016.txt", sep = ";")#, col.names = c("First", "Gender", "Count"))
colnames(dfNames) = c("Name", "Gender", "Count")
```

Lets look a summary of the data, and use the str funciton to see its strucutre and what the data looks like.
```{r}
#structure of the data
summary(dfNames)
str(dfNames)
```

There is a name with incorrect data. Noteably, the name contains three "y"s at the end of a name. Lets use RegEx to find it. Then remove it from the data frame. Save the results in a new object called "y2016".
```{r}
repeatName = dfNames[grep("yyy$", dfNames$Name),]
repeatName
```

The repeat name occure in row 212, so we will drop this observation
```{r}
repeatNum = which(dfNames$Name == repeatName$Name[1])
y2016 = dfNames[-212,]
```

## Question 2
Read in the names from 2015. The seperator has changed to a commma.
```{r}
y2015= read.table("yob2015.txt", sep = ",", col.names = c("Name", "Gender", "Count"))
```

Display the last 10 rows in the dataframe. What is interesting about the data?
```{r}
tail(y2015, 10)
```
What is interesting about this data is that they are all boys names that were give out exactly five times starting with the letter Z. Seems unlikely.

Now Merge the data to dataframe y2016 by name column.
```{r}
#lets see if there is any 0 count names
x = sapply(y2015$Count, is.nan)
paste0("there are ",sum(x), " names that had N/A data")

y = sapply(y2016$Count, is.nan)
paste0("there are ",sum(y), " names that had N/A data")

#Is there any missing gender variables
gen15 = sapply(y2015$Gender, is.na)
paste0("there are ",sum(gen15), " names that had N/A data")

gen16 = sapply(y2016$Gender, is.na)
paste0("there are ",sum(gen16), " names that had N/A data")


#merge the data
final = merge(y2015, y2016, by="Name")
#final[duplicated(final$Name),]
```

## Question 3

What are the total number of children given the names listed in the dataframe in the years 2015 and 2016?
```{r}
final["Total"] = final$Count.x + final$Count.y
#there are lots of repeat numbers each year... are the data sets the same? Does not look like it.
#final["same"] = final$Count.x == final$Count.y
#final = final[-7]
```

What are the 10 most popular names?
```{r}
ordered = final[order(final$Total, decreasing = TRUE),]
head(ordered,10)
```

What are the 10 most popular girl names?
```{r}
girls = final[final$Gender.x == "F" & final$Gender.y == "F",]
girl = head(girls[order(girls$Total, decreasing = TRUE),],10)
girl[c("Name", "Total")]
write.csv(girl[c("Name", "Total")], "itsagirl.csv", row.names = FALSE)
```


## Question 4

My work can be found at: https://github.com/khthomas/SMU_MSDS/tree/master/MSDS6306_DoingDS/homework/Unit5  
