---
title: "Unit 9 - Homework"
author: "Kyle Thomas"
date: "October 28, 2017"
output: html_document
---
# Harry Potter and the Deathly Hallows: Part 2
# Web Scrapping Exercise
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(xml2)
library(XML)
library(htmltools)
library(dplyr)
library(rvest)
library(selectr)
library(tidyr)
library(kableExtra)
library(knitr)
library(ggplot2)
```

I scrapped the webpage using rvest and put the results into a data frame. The data frame contained several 
columns that were not useful. The first column was simply empty, and the third column contained ellipses. 
These columns were dropped from the dataframe. The names of the columns were changed to Actor and Character.
Finally, the blank first row is removed from the dataframe.
```{r init}
link = "http://www.imdb.com/title/tt1201607/fullcredits?ref_=tt_ql_1"
page = read_html(link)
#allTables = html_nodes(page, "table")
#cast2 = html_table(allTables[3], fill=TRUE)

#experimenting with forward pipe
castlist = page %>%
  html_nodes("table") %>%
  .[3] %>%
  html_table(fill=TRUE)

#Using the forward pipe to make a dataframe 
Cast = castlist %>%
  data.frame() %>%
  .[-1,-c(1,3)] #drop the blank first row and the useless columns
  
#rename
names(Cast) = c("Actor","Character")

```

After examining the Cast Dataframe it became apparent that some the data frame needed cleaning. First, there 
was an observation called "Rest of the cast listed alphabetically". This row was removed. Also, Mr. Warwick's 
character line was cleaned up.

```{r cleaning}
#Charning Mr. Warwick's character
Cast[11,2] = "Griphook//Professor Filius Flitwick"

#remove "Rest of the cast listed alphabetically" row
Cast = Cast[-c(92:92),]

```

Split the actor's names into two columns (one for first and last name). Show the first 10 rows of final output.
```{r split_names}
Cast = separate(Cast, Actor, into=c("FirstName", "Surname"), sep="[ ](?=[^ ]+$)")

#I have some wierd problems with kable very similar to: #https://stackoverflow.com/questions/46820236/first-column-outside-of-kable-table-with-kable-styling
kable(head(Cast,10), "html", align="l", row.names = FALSE) #%>% 
  # column_spec(1, width="1in") %>% 
  # column_spec(2, width="1in") %>%
  # column_spec(3, width="3in")
```

# NBA Shooting Statistics Web Scrapping Exercise
```{r sportball}
#Get the data and figure out which node is needed
espnLink = "http://www.espn.com/nba/team/stats/_/name/sa/san-antonio-spurs"
espnhtml = read_html(espnLink)
sportsTable = html_nodes(espnhtml, "table")

#make the dataframe
dfshooting = sportsTable %>%
  .[2] %>%
  html_table(fill=TRUE) %>%
  data.frame()

#make the columns a meaningful name and drop out rows that are useless
names(dfshooting) = dfshooting[2,]
dfshooting = dfshooting[-c(1,2),]

#seperate names from position
dfshooting = separate(dfshooting, PLAYER, into=c("Name", "Position"), sep=", ")
n = names(dfshooting)

#what are the data types of each column?
sapply(dfshooting, class)

#change into float
dfshooting = data.frame(lapply(dfshooting, function(x) type.convert(as.character(x), as.is=TRUE)))
sapply(dfshooting, class)
#names(dfshooting) = n #rename again due to revious funciton stripping out names

#graph field goals percentage per game
ggplot(data=dfshooting, aes(x=Name, y=FG., group=Position)) + 
  geom_col(aes(color=Position, fill=Position)) + 
  coord_flip() + 
  labs(title="San Antonio Spurs: Field Goals Percentage Per Game", x="Player", y="Average Field Goal Percentage")


```





























